---
title: ''
date: '`r format(Sys.time(), "%d, %B, %Y")`'
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '6'
---

<!--
Copyright 2019 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(bcdata) #available from GitHub https://bcgov.github.io/bcdata/
library(janitor)
library(dplyr)
library(lubridate)
library(stringr)
library(ggplot2)
library(readr)
library(skimr)
```

# IM/IT Procurement Data: Exploration

An exploration of public IM/IT Procurement data for British Columbia.


### Raw Data
This analysis uses publicly available data from the B.C. Data Catalogue---ministry-contract-awards-province-of-british-columbia record---released under the [B.C. Crown Copyright Licence](https://www2.gov.bc.ca/gov/content?id=1AAACC9C65754E4D89A118B875E0FBDA)

Metadata record for ministry-contract-awards-province-of-british-columbia record:

```{r review-data}
## Look at ministry-contract-awards-province-of-british-columbia record 
## in the B.C. Data Catalogue
bcdc_get_record("9bff5da9-fced-4671-8ff3-f6117e5c8266")
# bcdc_browse("9bff5da9-fced-4671-8ff3-f6117e5c8266") #open in browser
```

Pull the data set and get familiar with the variables:

```{r get-data}
## Get the tabular data with bcdata package
## Data released under B.C. Crown Copyright Licence
## https://www2.gov.bc.ca/gov/content?id=1AAACC9C65754E4D89A118B875E0FBDA
# awards_raw <- bcdc_get_data(record = "9bff5da9-fced-4671-8ff3-f6117e5c8266",
#                         resource = "aee0fed5-e1f5-4e32-b17c-c98eafdd0d36")
# 
# if (!exists("tmp"))  dir.create('tmp', showWarnings = FALSE)
# awards_raw %>% write_rds("tmp/awards_raw.rds")

awards_raw <- read_rds("tmp/awards_raw.rds")

head(awards_raw)
```


### Tidy Data

Tidy the raw data frame, e.g. convert award amounts from text to numbers, make a year column,
drop columns not likely to be informative at this time:

```{r clean-data}
## Tidy raw data frame
awards_tidy <- awards_raw %>% 
  clean_names() %>%
  mutate(year = year(date_awarded),
         award_total = str_remove_all(award_total, "[,]"),
         award_total = as.numeric(str_remove(award_total, "[$]")))  %>%
  select(year, issued_for_organization, title, successful_vendor,
         successful_vendor_city, award_total, province, country)

skim(awards_tidy) %>% kable()
```

Find IM/IT awards and build IM/IT categories from text descriptors in the `title` field:
```{r imit-rows}

```


### Data Visualizations & Summaries

#### Total procurement awards per year (2012-2018):

```{r}
## Total Awarded by Year
awards_tidy %>% 
  group_by(year) %>% 
  summarise(total = sum(award_total)/1000000) %>% 
  ggplot() +
  geom_col(aes(x = year, y = total), alpha = 0.6, fill = "blue") +
  labs(y = "Total Awarded (million $)",
       x = "",
       title = "Total Procurement Awards per Year (2012-2018)",
       subtitle = "Only includes awards through BCBids",
       caption = "Data from the B.C. Data Catalogue") +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
```

