---
title: ''
date: '`r format(Sys.time(), "%d, %B, %Y")`'
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '6'
---

<!--
Copyright 2019 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

library(bcdata) #source data from the B.C. Data Catalogue
library(janitor) #clean_names()
library(dplyr) #data munging
library(lubridate) #working with dates
library(stringr) #working with strings
library(ggplot2) #plotting
library(readr) #read in csv files
library(readxl) #read in xlsx files
library(tidytext) #tidying text data
```

# IM/IT Payments Data: Exploration

An exploration of IM/IT Payments data for British Columbia. This document outlines the exploration phase---getting and cleaning bcgov payments data from the B.C. Data Catalogue, determining which payments are IM/IT-related, and some descriptive summaries of the data.


### Raw Data

#### Payments

Get the publicly available data from the B.C. Data Catalogue---Consolidated Revenue Fund Detailed Schedules of Payments - Other Supplier Payments record---released under the [Open Government Licence &mdash; British Columbia licence](https://www2.gov.bc.ca/gov/content?id=A519A56BC2BF44E4A008B33FCF527F61).

Read the metadata record for consolidated-revenue-fund-detailed-schedules-of-payments-other-supplier-payments record:

```{r review-bcdc-data}
## Look at consolidated-revenue-fund-detailed-schedules-of-payments-other-supplier-payments record 
## in the B.C. Data Catalogue
bcdc_get_record("f6bad1d5-d1dc-4fdd-a076-650f55cc3fbc")
# bcdc_browse("f6bad1d5-d1dc-4fdd-a076-650f55cc3fbc") #open in browser
```

Pull the B.C. Data Catalogue data set and get familiar with the variables:

```{r get-bcdc-data}
## Get the tabular data with bcdata package
## Data released under released under the released under the [Open Government Licence - British Columbia licence](https://www2.gov.bc.ca/gov/content?id=A519A56BC2BF44E4A008B33FCF527F61).
payments_raw <- bcdc_get_data(record = "f6bad1d5-d1dc-4fdd-a076-650f55cc3fbc",
                        resource = "54473f32-6ba5-4075-984a-745fef3c5806")

if (!exists("tmp"))  dir.create('tmp', showWarnings = FALSE)
payments_raw %>% write_rds("tmp/payments_raw.rds")


glimpse(payments_raw)
```



