---
title: ''
date: '`r format(Sys.time(), "%d, %B, %Y")`'
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '6'
---

<!--
Copyright 2019 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(bcdata) #available from GitHub https://bcgov.github.io/bcdata/
library(janitor)
library(dplyr)
library(lubridate)
library(stringr)
library(ggplot2)
library(readr)
library(skimr)
library(readxl)
```

# IM/IT Procurement Data: Exploration

An exploration of IM/IT Procurement data for British Columbia.


### Raw Data

Starting with the publicly available data from the B.C. Data Catalogue---ministry-contract-awards-province-of-british-columbia record---released under the [B.C. Crown Copyright Licence](https://www2.gov.bc.ca/gov/content?id=1AAACC9C65754E4D89A118B875E0FBDA).

Metadata record for ministry-contract-awards-province-of-british-columbia record:

```{r review-data}
## Look at ministry-contract-awards-province-of-british-columbia record 
## in the B.C. Data Catalogue
bcdc_get_record("9bff5da9-fced-4671-8ff3-f6117e5c8266")
# bcdc_browse("9bff5da9-fced-4671-8ff3-f6117e5c8266") #open in browser
```

Pull the B.C. Data Catalogue data set and get familiar with the variables:

```{r get-data}
## Get the tabular data with bcdata package
## Data released under B.C. Crown Copyright Licence
## https://www2.gov.bc.ca/gov/content?id=1AAACC9C65754E4D89A118B875E0FBDA
# awards_raw <- bcdc_get_data(record = "9bff5da9-fced-4671-8ff3-f6117e5c8266",
#                         resource = "aee0fed5-e1f5-4e32-b17c-c98eafdd0d36")
# 
# if (!exists("tmp"))  dir.create('tmp', showWarnings = FALSE)
# awards_raw %>% write_rds("tmp/awards_raw.rds")

awards_bcbid <- read_rds("tmp/awards_raw.rds")

head(awards_bcbid)
```

Next, pull in the program-area provided BCDevExchange code-with-us data set:

```{r}
awards_cwu <- read_xlsx(here::here("data/Code With Us Opportunity Tracking.xlsx"),
                 sheet = "Sheet1", skip = 6)

head(awards_cwu)
```


### Tidy Data

Tidy the raw data frames, e.g. convert award amounts from text to numbers, make a fiscal year column,
drop columns not likely to be informative at this time and so on:

```{r clean-data}

## Convert award date to fiscal year
convert_to_fy <- function(date){
fy <- as.integer(format(date, "%Y")) + (as.integer(format(date, "%m")) >= 4)
paste0(fy-1, "/", str_sub(fy, 3, 4))
}


## Tidy awards_bcbids raw data frame
awards_bcbid_tidy <- awards_bcbid %>% 
  clean_names() %>%
  mutate(fiscal_year = convert_to_fy(date = date_awarded),
         award_total = str_remove_all(award_total, "[,]"),
         award_total = as.numeric(str_remove(award_total, "[$]")))  %>%
  select(fiscal_year, issued_for_organization, title, successful_vendor,
         successful_vendor_city, award_total, province, country) %>% 
  mutate(data_source = "bcdc_bcbid")

skim(awards_bcbid_tidy ) %>% kable()


## Tidy awards_cwu raw data frame
awards_cwu_tidy <- awards_cwu %>% 
  clean_names() %>% 
  filter(!is.na(amount_paid),
         !is.na(assigned_to)) %>% 
  select(award_total = price, title, fiscal_year = fiscal_year_payment_was_made,
         issued_for_organization = ministry, successful_vendor = assigned_to) %>% 
  mutate(award_total = as.numeric(award_total),
         province = NA_character_,
         successful_vendor_city = NA_character_,
         country = NA_character_,
         data_source = "cwu")

## Join data sets
awards_tidy <- awards_bcbid_tidy %>%
  bind_rows(awards_cwu_tidy)

```

Find IM/IT awards and build IM/IT categories from text descriptors in the `title` field:
```{r imit-rows}

```


### Data Visualizations & Summaries

#### Total procurement awards per fiscal year (2012-2018):

```{r}
## Total Awarded by Year
awards_tidy %>% 
  filter(fiscal_year != c("2011/12"),
         fiscal_year != c("2019/20")) %>% 
  group_by(fiscal_year) %>% 
  summarise(total = sum(award_total)/1000000) %>% 
  ggplot() +
  geom_col(aes(x = fiscal_year, y = total), alpha = 0.6, fill = "blue") +
  labs(y = "Total Awarded (million $)",
       x = "",
       title = "Total Procurement Awards per Year (2012-2018)",
       subtitle = "Only includes awards through BCBids & BCDevExchange Code With Us",
       caption = "Data from the B.C. Data Catalogue & BCDevExchange") +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
```

