---
title: ''
date: '`r format(Sys.time(), "%d, %B, %Y")`'
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '6'
---

<!--
Copyright 2019 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

library(bcdata) #source data from the B.C. Data Catalogue
library(purrr)  #map over resources for reading in data
library(dplyr) #data munging
library(stringr) #working with strings
library(readr) #read/write files
library(ggplot2) #plotting
library(scales) #comma
# library(tidytext) #tidying text data
```

# IM/IT Payments Data: Exploration

An exploration of IM/IT Payments data for British Columbia. This document outlines the exploration phase---getting and cleaning bcgov payments data from the B.C. Data Catalogue, determining which payments are IM/IT-related, and some descriptive summaries of the data.


### Raw Data

#### Payments

Get the publicly available data from the B.C. Data Catalogue---Consolidated Revenue Fund Detailed Schedules of Payments - Other Supplier Payments record---released under the [Open Government Licence &mdash; British Columbia licence](https://www2.gov.bc.ca/gov/content?id=A519A56BC2BF44E4A008B33FCF527F61): <https://catalogue.data.gov.bc.ca/dataset/consolidated-revenue-fund-detailed-schedules-of-payments-other-supplier-payments>

Get the raw data:

```{r get-bcdc-data}
## Open the record in default browser
# bcdc_browse("f6bad1d5-d1dc-4fdd-a076-650f55cc3fbc") #open in browser

## Get record for consolidated-revenue-fund-detailed-schedules-of-payments-other-supplier-payments  
## data in the B.C. Data Catalogue.
## Data released under released under the released under the [Open Government Licence - British Columbia licence](https://www2.gov.bc.ca/gov/content?id=A519A56BC2BF44E4A008B33FCF527F61)
payment_record <- bcdc_get_record("f6bad1d5-d1dc-4fdd-a076-650f55cc3fbc")

## Map over and merge the tabular data
# map_chr(payment_record$resources, "id")

payments_resources_df <- payment_record$resource_df %>% 
  mutate(year = str_sub(name,-31, -27))

# df2017 <- bcdc_get_data("f6bad1d5-d1dc-4fdd-a076-650f55cc3fbc", "237a7c70-7acb-4e89-be64-23c2fec61f0d")

payments_df_list <- map2(.x = payments_resources_df$id, 
        .y = payments_resources_df$year, 
        .f = ~{bcdc_get_data(record = "f6bad1d5-d1dc-4fdd-a076-650f55cc3fbc",
                        resource = .x) %>%
         mutate(year = as.numeric(.y))
          })

map(payments_df_list, ~ colnames(.x))

# 2020 file has metadata at the top, needs to be cleaned for the code to run

# payments_df_list[[1]]  <- payments_df_list[[1]] %>%


payments_raw <- map2_dfr(.x = payments_resources_df$id, 
        .y = payments_resources_df$year, 
        .f = ~{bcdc_get_data(record = "f6bad1d5-d1dc-4fdd-a076-650f55cc3fbc",
                        resource = .x) %>%
         select(1:2) %>%
            rename(name = 1,          #renaming columns using position, 
                   amount = 2) %>%    #fragile, but names differ across dfs
         mutate(year = as.numeric(.y))
          })

if (!exists("tmp"))  dir.create('tmp', showWarnings = FALSE)
payments_raw %>% write_rds("tmp/payments_raw.rds")

glimpse(payments_raw)
```



```{r imit-rows}

# payments_raw %>% filter(str_detect(name, "IBM")) %>% distinct(name)

payments_imit <- payments_raw %>%
  mutate(imit = case_when(str_detect(name, "IBM") ~ "IBM",
                          str_detect(name, "TELUS") ~ "Telus",
                          str_detect(name, "CGI GROUP") ~ "CGI",
                          str_detect(name, "CGI INFORMATION") ~ "CGI",
                          str_detect(name, "ORACLE") ~ "Oracle/Sun",
                          str_detect(name, "SUN MICROSYSTEMS") ~ "Oracle/Sun",
                          str_detect(name, "PACKARD") ~ "Hewlett-Packard/EDS/ESIT",
                          str_detect(name, "EDS CANADA") ~ "Hewlett-Packard/EDS/ESIT",
                          str_detect(name, "EDS ADVANCED") ~ "Hewlett-Packard/EDS/ESIT",
                          str_detect(name, "ESIT CANADA") ~ "Hewlett-Packard/EDS/ESIT",
                          str_detect(name, "ESIT ADVANCED") ~ "Hewlett-Packard/EDS/ESIT",
                          str_detect(name, "DELOITTE") ~ "Deloitte",
                              TRUE ~ NA_character_))

glimpse(payments_imit)
```



```{r}
## Payment Totals by Year Over Time
payments_imit %>% 
  filter(!is.na(imit)) %>% 
  group_by(year, imit) %>% 
  summarise(total = sum(amount)/1000000) %>% 
  ggplot() +
  geom_area(aes(x = year, y = total, fill = imit), alpha = 0.5) +
  scale_y_continuous(expand = c(0,0)) +
  # limits = c(0, 150), breaks = seq(0, 150, 50),
  scale_x_continuous(limits = c(2007, 2019), 
                     breaks = seq(2008, 2019, 1), 
                     expand = c(0,0)) +
  labs(y = "Total Payments (Million $)",
       x = "",
       fill = "",
       title = "IMIT Payments (2007-2019)",
       caption = "Data from the B.C. Data Catalogue") +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
```
