---
title: ''
date: '`r format(Sys.time(), "%d, %B, %Y")`'
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '6'
---

<!--
Copyright 2019 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

library(bcdata) #available from GitHub https://bcgov.github.io/bcdata/
library(janitor) #clean_names()
library(dplyr) #data munging
library(lubridate) #working with dates
library(stringr) #working with strings
library(ggplot2) #plotting
library(readr) #read in csv files
library(readxl) #read in xlsx files
library(tidytext) #tidying text data
```

# IM/IT Procurement Data: Exploration

An exploration of IM/IT Procurement data for British Columbia. This document outlines the exploration phase---finding, getting and cleaning multiple, different data sources, determining which awards are IM/IT-related, and some descriptive summaries of the combined data.


### Raw Data

#### BCBid

Get the publicly available data from the B.C. Data Catalogue---ministry-contract-awards-province-of-british-columbia record---released under the [B.C. Crown Copyright Licence](https://www2.gov.bc.ca/gov/content?id=1AAACC9C65754E4D89A118B875E0FBDA).

Read the metadata record for ministry-contract-awards-province-of-british-columbia record:

```{r review-bcdc-data}
## Look at ministry-contract-awards-province-of-british-columbia record 
## in the B.C. Data Catalogue
bcdc_get_record("9bff5da9-fced-4671-8ff3-f6117e5c8266")
# bcdc_browse("9bff5da9-fced-4671-8ff3-f6117e5c8266") #open in browser
```

Pull the B.C. Data Catalogue data set and get familiar with the variables:

```{r get-bcdc-data}
## Get the tabular data with bcdata package
## Data released under B.C. Crown Copyright Licence
## https://www2.gov.bc.ca/gov/content?id=1AAACC9C65754E4D89A118B875E0FBDA
# awards_raw <- bcdc_get_data(record = "9bff5da9-fced-4671-8ff3-f6117e5c8266",
#                         resource = "aee0fed5-e1f5-4e32-b17c-c98eafdd0d36")
# 
# if (!exists("tmp"))  dir.create('tmp', showWarnings = FALSE)
# awards_raw %>% write_rds("tmp/awards_raw.rds")

awards_bcbid <- read_rds("tmp/awards_raw.rds")

glimpse(awards_bcbid)
```


#### Code-With-Us

Next, pull in the program-area provided BCDevExchange code-with-us data set:

```{r get-cwu-data}
awards_cwu <- read_xlsx(here::here("data/Code With Us Opportunity Tracking.xlsx"), sheet = "Sheet1", skip = 6) %>% 
  select(-EMAIL, -`CONTACT NAME`)

glimpse(awards_cwu)
```


#### Open Information Webpage

Data on contracts over $10,000 are available on-line, housed in PDF and XLS files (with different formats over time)---by Ministry for each quarter from April 2017 to June 2019: [https://www2.gov.bc.ca/gov/content?id=AFBE378C5C684CF980CD6008364BD868](https://www2.gov.bc.ca/gov/content?id=AFBE378C5C684CF980CD6008364BD868)


Data on directly awarded contracts are available on-line, housed in PDF and XLS files (with different formats over time)---available by Ministry for each month from 2016-17 to June 2019: [https://www2.gov.bc.ca/gov/content?id=217BF2AE401E44F1A5DA3CB4C36E95AF](https://www2.gov.bc.ca/gov/content?id=217BF2AE401E44F1A5DA3CB4C36E95AF)

_**NOTE:**_ None of the data sets include information on award type (e.g. IM/IT) so users need to determine this from the available fields.

### Tidy Data

Tidy the multiple raw data frames, e.g. convert award amounts from text to numbers, make a fiscal year column, drop columns not likely to be informative at this time and so on, and bind together into one data frame:

```{r clean-data}
## Convert award date to fiscal year
convert_to_fy <- function(date){
fy <- as.integer(format(date, "%Y")) + (as.integer(format(date, "%m")) >= 4)
paste0(fy-1, "/", str_sub(fy, 3, 4))
}


## Tidy awards_bcbids raw data frame
awards_bcbid_tidy <- awards_bcbid %>% 
  clean_names() %>%
  mutate(fiscal_year = convert_to_fy(date = date_awarded),
         award_total = str_remove_all(award_total, "[,]"),
         award_total = as.numeric(str_remove(award_total, "[$]")))  %>%
  select(fiscal_year, issued_for_organization, title, successful_vendor,
         successful_vendor_city, award_total, province, country) %>% 
  mutate(data_source = "bcdc_bcbid")


## Tidy awards_cwu raw data frame
awards_cwu_tidy <- awards_cwu %>% 
  clean_names() %>% 
  filter(!is.na(amount_paid),
         !is.na(assigned_to)) %>% 
  select(award_total = price, title, fiscal_year = fiscal_year_payment_was_made,
         issued_for_organization = ministry, successful_vendor = assigned_to) %>% 
  mutate(award_total = as.numeric(award_total),
         province = NA_character_,
         successful_vendor_city = NA_character_,
         country = NA_character_,
         data_source = "cwu")

## Join data sets
awards_tidy <- awards_bcbid_tidy %>%
  bind_rows(awards_cwu_tidy)

glimpse(awards_tidy)
```


### Find IM/IT Awards

Find IM/IT awards and build IM/IT categories from text descriptors in the `title` field.

Look at the most common words (n > 100) in the `title` field, as a start:

```{r title-text}
title_text <- awards_tidy %>%
  unnest_tokens(word, title) %>%
  anti_join(stop_words) %>% 
  filter(str_detect(word, "^[a-zA-Z]+"))


title_text %>%
  count(word, sort = TRUE) %>% 
  filter(n > 100) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) +
  geom_col(alpha = 0.6) +
  xlab(NULL) +
  coord_flip() +
  theme_minimal()


# title_text %>%
#     count(issued_for_organization, word, sort = TRUE) %>% 
#     filter(n > 100) %>%
#     mutate(word = reorder(word, n)) %>%
#     ggplot(aes(word, n)) +
#     geom_col() +
#     facet_wrap(~issued_for_organization, ncol = 2) +
#     xlab(NULL) +
#     coord_flip() +
#     theme_minimal()
```

Go through `title` descriptions and assign into categories based on data source or presence/absence of im/it-related words (e.g. `software` `==` `imit-software`):

```{r imit-rows}
imit_words <- c("software", "data", "licenses", "information", "technology",
                "licence", "license")

awards_imit <- awards_tidy %>%
  mutate(title = str_to_lower(title),
         subcategory = case_when(str_detect(title, "software") ~ "imit-software",
                              str_detect(title, "license") ~ "imit-software",
                              str_detect(title, "licence") ~ "imit-software",
                              str_detect(title, "network") ~ "imit-hardware",
                              str_detect(title, "laptop") ~ "imit-hardware",
                              str_detect(title, "architect support") ~ "imit-support",
                              data_source == "cwu" ~ "imit",
                              TRUE ~ "everything else"),
         category = case_when(str_detect(subcategory, "imit") ~ "imit",
                              TRUE ~ "everything else"))

glimpse(awards_imit)
```


_**NEXT**_ Work with IM/IT subject experts to identify a representative sample of awards that are IM/IT and IM/IT subcategories to train a predictive text model. 


### Descriptive Summaries & Visualizations

Summarize total amount awarded by category by fiscal year:

```{r}
## Total Amount Awarded by Year
awards_imit %>% 
  filter(fiscal_year != c("2011/12"),
         fiscal_year != c("2019/20")) %>% 
  group_by(fiscal_year, category) %>% 
  summarise(total = sum(award_total)/1000000) %>% 
  ggplot() +
  geom_col(aes(x = fiscal_year, y = total, fill = category), alpha = 0.5) +
  scale_fill_manual(name="", values = c("imit" = "blue", "everything else" = "grey60")) +
  labs(y = "Total Amount Awarded (million $)",
       x = "",
       title = "Total Amount Awarded per Year (2012/13-2018/19)",
       subtitle = "IM/IT category only includes code-with-us & awards related to software at this time",
       caption = "Data from the B.C. Data Catalogue & BCDevExchange") +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
```

